Intelligence and Age at First Intercourse: Cause or Confound?

### 2016-09-24 Common/content/writing/results_within.tex
update

We replicated the between-family analyses reported in the previous subsection, adding within-family difference scores as independent variables to the between-family means. Using the discordant sibling model, we predicted the differences in Generation 2 AFI as a function of differences in intelligence, controlling for means of the outcomes and predictors. We ran three series of models, where we examined the individual and then joint influence of Gen1 intelligence and Gen2 intelligence. Moreover, within each series we included three Generation 2 linking method variants, just as we did in the between family analyses: the mixed model reports the differences of the firstborns of each sister, the daughters model reports the differences of the firstborn girls, and the sons model reports the differences of the firstborn sons. In the spirit of transparency, we have reported all three methods. However, because all three linking methods resulted in similar findings, we will focus the results section on the mixed model and only discuss the other two methods when they deviate.

@@

Moreover, when we replicate this analysis using the larger Gen1 sample of siblings, we find that the between-family relationship of Gen1 AFI and Gen1 intelligence is moderate $(r =.291)$, but within the family, the AFI of the smarter sibling was no later than the less smart sibling's. In Figure \ref{plot_gen1_afi}, we have illustrated this finding using a scatter plot for the individual subjects, and two marginal distributions to show the mean levels of AFQT and AFI, grouped by sibling classification. The distribution above the x-axis shows the distributions of AFQT based on sibling classification. The blue distribution shows the standardized AFQT score for the siblings who were .33 standard-deviations smarter than their sibling, the red distribution shows the standardized AFQT score of siblings who were at least .33 standard deviations less smart than their sibling, and the purple distribution shows those siblings who were within at least .33 standard deviations of one another. As expected, the blue distribution of smarter siblings had a higher mean than the other two groups. The distribution across from the y-axis shows the distribution of AFI based on those same sibling classifications. All three of these distributions are indistinguishable, which suggests that siblings do not differ in their AFI across these intelligence categories. This plot illustrate that there is no within-family effect for intelligence on AFI. Had there been a within-family effect of intelligence on AFI, the marginal distributions of AFI would be visually separable, and would follow the pattern or the separate distributions of AFQT.%\pagebreak

>>>>

We replicated the between-family analyses reported in the previous subsection, adding within-family difference scores as independent variables to the between-family means. Using the discordant sibling model, we predicted the differences in Generation 2 AFI as a function of differences in intelligence, controlling for means of the outcomes and predictors. We ran three series of models, where we examined the individual and then joint influence of Gen1 intelligence and Gen2 intelligence. Moreover, within each series we included three Generation 2 linking method variants, just as we did in the between family analyses: the mixed model reports the differences of the firstborns of each sister, the daughters model reports the differences of the firstborn girls, and the sons model reports the differences of the firstborn sons. In the spirit of transparency, we have reported all three methods. However, because all three linking methods resulted in similar findings, we will focus the results section on the mixed model and only discuss the other two methods when they deviate. We have also provided zero-order and pairwise semi-partial correlations for all between-family variables (Tables \ref{table_cor} and \ref{table_spcor_wtn}), using the mixed-model data and the ppcor \R library \citep{kim2015ppcor}.

@@

Moreover, when we replicate this analysis using the larger Gen1 sample of siblings, we find that the between-family relationship of Gen1 AFI and Gen1 intelligence is moderate $(r =.291)$, but within the family, the AFI of the smarter sibling was no later than the less smart sibling's. In Figure \ref{plot_gen1_afi}, we have illustrated this finding using a scatter plot for the individual subjects, and two marginal distributions to show the mean levels of AFQT and AFI, grouped by sibling classification. The distribution above the x-axis shows the distributions of AFQT based on sibling classification. The blue distribution shows the standardized AFQT score for the siblings who were .33 standard-deviations smarter than their sibling, the red distribution shows the standardized AFQT score of siblings who were at least .33 standard deviations less smart than their sibling, and the purple distribution shows those siblings who were within at least .33 standard deviations of one another. As expected, the blue distribution of smarter siblings had a higher mean than the other groups. The distribution across from the y-axis shows the distribution of AFI based on those same sibling classifications. All three of these distributions are indistinguishable, which suggests that siblings do not differ in their AFI across these intelligence categories. This plot illustrate that there is no within-family effect for intelligence on AFI. Had there been a within-family effect of intelligence on AFI, the marginal distributions of AFI would be visually separable, and would follow the pattern for the separate distributions of AFQT.%\pagebreak

### 2016-09-24 Common/content/writing/subjects.tex
update

In 1986, all biological children of the female NLSY79 participants were surveyed, the first round of the NLSY79 Children and Young Adults (NLSY-Children) survey. Now that childbearing is complete among the NLSY79 females (who were 49 to 57 in the 2014 survey), a total of 11,512 respondents are in the NLSY-Children surveys, which is continuing on a biennial basis. Participants in the NLSY79 will typically be referred to as the Generation 1(Gen1) sample, whereas the NLSY-Children will be referred to as the Generation 2(Gen2) sample.

>>>>

In 1986, all biological children of the female NLSY79 participants were surveyed for the first round of the NLSY79 Children and Young Adults (NLSY-Children) survey. Now that childbearing is complete among the NLSY79 females (who were 49 to 57 in the 2014 survey), a total of 11,512 respondents are in the NLSY-Children surveys, which is continuing on a biennial basis. Participants in the NLSY79 will typically be referred to as the Generation 1 (Gen1) sample, whereas the NLSY-Children will be referred to as the Generation 2 (Gen2) sample.
